Designing and modeling a browser-based distributed evolutionary computation systems

### 2015-04-01 modeling-volunteer-computing-gecco.tex
Completing biblio
Expanding server description

The server, that was also introduced during \cite{arxiv:modeling}, has
been rewritten in part to add new features that make experimentation
easier. I has been written using node.js and the {\tt express.js}
module and includes the following REST \cite{Castillo12REST} 
routes:
\item {\tt GET random} returns a random, non-evaluated individual,
  from the pool. The pool is just an array containing chromosomes,
  that is initialized at the beginning of the experiment.
\item {\tt PUT one} puts a single individual along with its fitness in the pool. In REST
  conventions, {\tt PUT} is used to create a resource, that is why it
  is used, in this case, to add a new individual to the pool.
\end{itemize}

@@

The server program includes more logic to serve the static pages
that contain the talk in which it is included and the JavaScript/CSS
code also, but it all amounts to a small amount of code. This code can
be deployed, in principle, in any server running node.js, but for the
purposes of this experiment Heroku, a Platform as a Service with a
free tier, has been chosen. It could be also be deployed, with small
or no modification, to OpenShift or other similar PaaS. Far from being
a freeriding way, using free or very low cost servers can be a
sustainable way of performing massive distributed evolutionary
computation experiments \cite{Merelo2014,jj:idc:lowcost}.

Having the pages and the server in the same domain allows to work in
default Ajax mode, that does not allow cross-site requests. However,
these can also be disabled if needed. In this case, every domain holds
a single experiment, there is no multi-tenancy of experiments, so its
management is quite simple.

>>>>

The server, that was also introduced during \cite{arxiv:modeling}, has
been rewritten in part to add new features that make experimentation
easier. I has been written using node.js and the {\tt express.js} and {\tt winston}
modules for REST and logging and includes the following REST \cite{Castillo12REST} 
routes:
\begin{itemize}
\item {\tt GET random} returns a random, non-evaluated individual,
  from the pool. The pool is just an array containing chromosomes,
  that is initialized at the beginning of the experiment.
\item {\tt PUT one} puts a single individual along with its fitness in the pool. In REST
  conventions, {\tt PUT} is used to create a resource, that is why it
  is used, in this case, to add a new individual to the pool.
\end{itemize}

@@

The server program includes also additional logic for logging to
different devices, routes to obtain information from the server and code to serve the static pages
and the JavaScript/CSS
code also. Additionally to the code in the previous paper
\cite{arxiv:modeling} this version restarts the chromosome pool when
the experiment has finished; it includes a configurable termination
condition that checks for the solution and empties the chromosome pool
and logs the end of the experiment. This allows us to run continuously
experiments without restarting the server by hand as we did in the
previous version.

All in all, it amounts to a few hundred lines of code
including all languages. This code can
be deployed, in principle, in any server that is able to run node.js, but for the
purposes of this experiment Heroku, a Platform as a Service with a
free tier, has been chosen. It could be also be deployed, with small
or no modification, to OpenShift or other similar PaaS. Far from being
a cheap way of performing experiments, using free or very low cost servers can be a
sustainable way of performing massive distributed evolutionary
computation experiments \cite{Merelo2014,jj:idc:lowcost}.

Having the pages and the server in the same domain allows to work in the
default Ajax mode, that does not allow cross-site requests. However,
these can also be enabled if needed so that the static pages and the
REST server can be hosted in different servers. In this case, every domain holds
a single experiment, there is no multi-tenancy of experiments, so its
management is quite simple.
